---
title: "Index"
author: "Rosie Hohnen"
date: "2026-03-19"
output: html_document
---
# OBP releases at low occupancy sites analysis

This file demonstrates the code and analysis undertaken to examine the drivers of survival for OBPs released at low occupancy sites between 1994 and 2022. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in libraries

```{r libraries, message=FALSE, warning=FALSE}
#----Load libraries and check data----
library(ggpubr)
library(PerformanceAnalytics)
library(tidyverse)
library(stats)
library(MuMIn)
library(lme4)
library(glmmTMB)
library(DHARMa)
library(ggplot2)
library(broom.mixed)
library(car)
library(performance)
library(patchwork)
library(ggplot2)
```

## Read in data

Also standardize variables and check correlations

```{r data, echo=TRUE, warning=FALSE}
# Read in script
Trans_2 <-read.csv('Transloc_sum_V5.csv', header = TRUE)
summary(Trans_2)

# Standardize key variables
Trans_2 <- Trans_2 %>%
  mutate(Transmitters_std = as.numeric(scale(Tranmitters))) %>%
  mutate(Training_std = as.numeric(scale(Training))) %>%
  mutate(Conspecifics_std = as.numeric(scale(Conspecifics))) %>%
  mutate(Total_release_std = as.numeric(scale(Total_release)))

# Check correlations
correlation_subset <- Trans_2[, c(10,12,14,15,30,31,32,33)]
chart.Correlation(correlation_subset, histogram=TRUE)
```

# Run models on translocation persistence

```{r models1, echo=TRUE}
#----Run models on translocation survival----
f1 <- glm(cbind(Transloc, Total_r - Transloc) ~ Total_release_std, family="binomial", data=Trans_2)
f2 <- glm(cbind(Transloc, Total_r - Transloc) ~ Transmitters_std, family="binomial", data=Trans_2)
f3 <- glm(cbind(Transloc, Total_r - Transloc) ~ Training_std, family="binomial", data=Trans_2)
f4 <- glm(cbind(Transloc, Total_r - Transloc) ~ Total_release_std + Transmitters_std + Training_std , family="binomial", data=Trans_2)
f5 <- glm(cbind(Transloc, Total_r - Transloc) ~ 1 , family="binomial", data=Trans_2)

summary(f4)
summary(f1)
```

# Run model selection

```{r model1, echo=TRUE}
# Run model selection
fmList1<-model.sel(f1=f1, f2=f2, f3=f3, f4=f4, f5=f5)
fmList1
```

# Check residuals

Most of the red flags appear to be derived from outliers which is hard to avoid with a small data set. The values of dispersion, uniformity are acceptable.
```{r resid1, echo=TRUE}
# Check residuals
sim_f4 <- simulateResiduals(fittedModel = f4, n = 1000)
plot(sim_f4)
testDispersion(sim_f4)
testUniformity(sim_f4)
testOutliers(sim_f4)

```

# Check VIF values

```{r vif1, echo=TRUE}
vif(f4)
```

# Run models on season persistence

```{r model2, echo=TRUE}
#----Run models on season survival----
b1 <- glm(cbind(Season_survive, Total_r - Season_survive) ~ Total_release_std, family="binomial", data=Trans_2)
b2 <- glm(cbind(Season_survive, Total_r - Season_survive) ~ Transmitters_std, family="binomial", data=Trans_2)
b3 <- glm(cbind(Season_survive, Total_r - Season_survive) ~ Training_std, family="binomial", data=Trans_2)
b4 <- glm(cbind(Season_survive, Total_r - Season_survive) ~ Total_release_std + Transmitters_std + Training_std , family="binomial", data=Trans_2)
b5 <- glm(cbind(Season_survive, Total_r - Season_survive) ~ 1 , family="binomial", data=Trans_2)

summary(b2)
summary(b4)
```

# Model selection of season survival models

```{r selection2, echo=TRUE}
# Run model selection
fmList<-model.sel(b1=b1, b2=b2, b3=b3, b4=b4, b5=b5)
fmList
```

# Check residuals

Again this looks mostly ok
```{r residuals2, echo=TRUE}
# Check residuals
sim_b2 <- simulateResiduals(fittedModel = b2, n = 1000)
plot(sim_b2)
testDispersion(sim_b2)
testUniformity(sim_b2)
testOutliers(sim_b2)
```

# Check vifs
```{r vifs, echo=TRUE}
vif(b4)
```

# Make model prediction plot

```{r modelpreds, echo=TRUE}
newdata <- data.frame(
  Total_release_std = seq(min(Trans_2$Total_release_std),
                          max(Trans_2$Total_release_std),
                          length = 100),
  Transmitters_std = mean(Trans_2$Transmitters_std, na.rm = TRUE),
  Training_std = mean(Trans_2$Training_std, na.rm = TRUE))

pred <- predict(f4, newdata = newdata, type = "response", se.fit = TRUE)

newdata$fit <- pred$fit
newdata$se <- pred$se.fit
newdata$lwr <- newdata$fit - 1.96 * newdata$se
newdata$upr <- newdata$fit + 1.96 * newdata$se
Trans_2$prop_survival <- Trans_2$Transloc / Trans_2$Total_r

f4plot <- ggplot() +
  geom_point(data = Trans_2,
             aes(x = Total_release_std, y = prop_survival),
             size = 2, alpha = 0.7) +
  geom_line(data = newdata,
            aes(x = Total_release_std, y = fit),
            size = 1.2, colour = "black") +
  geom_ribbon(data = newdata,
              aes(x = Total_release_std, ymin = lwr, ymax = upr),
              alpha = 0.2, fill = "black") +
  labs(x = "Release group size (standardized)",
       y = "Translocation survival probability") +
  theme_classic()

# Make model predictions plot
newdata <- data.frame(
  Transmitters_std = seq(min(Trans_2$Transmitters_std),
                         max(Trans_2$Transmitters_std),
                         length = 100))

pred <- predict(b2, newdata = newdata, type = "response", se.fit = TRUE)

newdata$fit <- pred$fit
newdata$se <- pred$se.fit
newdata$lwr <- newdata$fit - 1.96 * newdata$se
newdata$upr <- newdata$fit + 1.96 * newdata$se
Trans_2$prop_season_survival <- Trans_2$Season_survive / Trans_2$Total_r
b2plot <- ggplot() +
  geom_point(data = Trans_2,
             aes(x = Transmitters_std, y = prop_season_survival),
             size = 2, alpha = 0.7) +
  geom_line(data = newdata,
            aes(x = Transmitters_std, y = fit),
            colour = "black", linewidth = 1.2) +
  geom_ribbon(data = newdata,
              aes(x = Transmitters_std, ymin = lwr, ymax = upr),
              fill = "black", alpha = 0.2) +
  labs(x = "Transmitters (standardized)",
       y = "Predicted seasonal survival probability") +
  theme_classic()

f4plot + b2plot +
  plot_annotation(tag_levels = "a")
```

# Plot coef estimates

```{r coefest, echo=TRUE}
## Extract coefficient estimates
f4_coefs <- tidy(f4, effects = "fixed") %>%
  mutate(model = "f4")
b2_coefs <- tidy(b2, effects = "fixed") %>%
  mutate(model = "b2")
coef_data <- bind_rows(f4_coefs, b2_coefs)

coef_data <- coef_data %>%
  rename(Variable = term) %>%
  rename(Coefficent_estimate = estimate) %>%
  mutate(model = dplyr::recode(model, "f4" = "Translocation_survival")) %>%
  mutate(model = dplyr::recode(model, "b2" = "Season_survival"))


p <- ggplot(coef_data, aes(x = Coefficent_estimate, y = Variable)) +
  geom_point() +
  geom_errorbarh(aes(xmin = Coefficent_estimate - 1.96*std.error,
                     xmax = Coefficent_estimate + 1.96*std.error),
                 height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~model) +
  theme_bw()
p
```

# Make survival plots

```{r survival, echo=TRUE}
# Make survival plot

p1 <- ggplot(Trans_2,aes(x=Prop_Season_survival, y=Prop_Transloc)) +
  geom_point() +
  theme(panel.background = element_rect(fill = 'white', colour = 'darkgrey')) +
  geom_smooth(method=lm, level=0.99, col='black') +
  theme(panel.grid.major = element_line(colour = 'white')) +
  labs(x= "Proportion of birds that survive the season", y= "Proportion of birds that survive the translocation") +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black")) +
  theme(text = element_text(size = 12))  

p2 <- ggplot(Trans_2,aes(x=Prop_Returns,y=Prop_Season_survival)) +   geom_smooth(method=lm, level=0.99, col='black') +
  geom_point() +
  theme(panel.background = element_rect(fill = 'white', colour = 'darkgrey')) +
  theme(panel.grid.major = element_line(colour = 'white')) +
  labs(x= "Proportion of birds that return the following season", y= "Proportion of birds that survive the season") +
  theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black")) +
  theme(text = element_text(size = 12))  

p1 + p2 +
  plot_annotation(tag_levels = "a")
```

# Plot survival differences

```{r survival2, echo=TRUE}
## Plot survival differences between locations
Ave <-read.csv('Average_success2.csv', header = TRUE)
head(Ave)
p = ggplot(data=Ave, aes(x=Location, y=Percentage, fill = Value)) + geom_bar(stat="identity", position="dodge")
p = p + scale_color_manual(values=c("#333333", "#999999", "#CCCCCC")) + scale_fill_manual(values=c("#333333","#999999", "#CCCCCC"))
p = p + geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,position=position_dodge(.9))
p = p + theme(panel.background = element_rect(fill = 'white', colour = 'white'))
p = p + theme(panel.border = element_blank(), axis.line = element_line())
p = p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p = p + labs(x= "Location", y= "Average percentage")
p = p + theme(axis.text.x = element_text(colour = "black", size=12, margin= margin(t=0,r=20,b=0,l=0)), 
              axis.text.y = element_text(colour = "black", size=12, margin = margin(t=20,r=0,b=20,l=0)))
p = p + theme(axis.title=element_text(size=12))
p = p + theme(legend.text=element_text(size=12))
p = p + theme(legend.title=element_text(size=12))
print(p)
```

# Plot data source consistency

```{r consistency, echo=TRUE}
Ave <-read.csv('Consistency.csv', header = TRUE)
head(Ave)
p = ggplot(data=Ave, aes(x=Year, y=Birds_released, fill = Source)) + geom_bar(stat="identity", position="dodge")
#p = p + scale_color_manual(values=c("#333333", "#999999", "#CCCCCC")) + scale_fill_manual(values=c("#333333","#999999", "#CCCCCC"))
#p = p + geom_errorbar(aes(ymin=Lower, ymax=Upper), width=.2,position=position_dodge(.9))
p = p + theme(panel.background = element_rect(fill = 'white', colour = 'white'))
p = p + theme(panel.border = element_blank(), axis.line = element_line())
#p = p + scale_x_continuous(breaks = 0:11,
#labels = paste0(c("1994", "1996", "1999", "2000", "2001", "2002", "2003"), "2004", "2005", "2007", "2008", "2009"), "year")
#p = p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p = p + labs(x= "Year", y= "Number of birds released")
p = p + theme(axis.text.x = element_text(colour = "black", size=15, margin= margin(t=0,r=20,b=0,l=0)), 
              axis.text.y = element_text(colour = "black", size=15, margin = margin(t=20,r=0,b=20,l=0)))
p = p + theme(axis.title=element_text(size=12))
p = p + scale_fill_brewer(palette="YlGnBu")
p = p + theme(legend.text=element_text(size=12))
p = p + theme(legend.title=element_text(size=12))
print(p)
```